\vspace{1.0cm} \Large \bf Supplemental material of ``Health and disease imprinted in the time variability of the human microbiome''

### 2017-01-12 mSystems/rev/mSys_discussion_rev.tex
Corrections and pre-submission, which leads to REVISION draft I
Changes to be committed:
	modified:   figs/supfig_taxlev1.eps
	modified:   figs/supfig_taxlev2.eps
	modified:   mSys_MAIN_rev.pdf
	modified:   mSys_MAIN_rev.tex
	modified:   mSys_MAIN_rev_marked.docx
	modified:   mSys_MandM_rev.tex
	modified:   mSys_discussion_rev.tex
	modified:   mSys_suppl_rev.pdf
	modified:   mSys_suppl_rev.tex
	new file:   tab/mSys_suppl_tab1.pdf
	new file:   tab/mSys_suppl_tab2.pdf
	new file:   tab/mSys_suppl_tab3.pdf
	new file:   tab/mSys_suppl_tab4.pdf
	new file:   tab/mSys_suppl_tab5.pdf
	new file:   tab/m

By solving Langevin's differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability into one of two phases, according to the ranking stability of the system. As we can see in the phase--space in Figure \ref{fig:main3}, three different conditions that could occur are shown. First, we could have a healthy microbiota with some fluctuations, as shown by one of the subjects of Caporaso \emph{et al.}'s study \cite{moving}. Because this case would have good fitness, its temporal variability would not place the microbiota in the unstable phase of the diagram. Second, we have a subject from the study by Dethlefsen and Relman \cite{antibiotic} who was perturbed twice by an antibiotic intake. His microbiota underwent sufficient change so as to lose its stability, and hence be placed in the unstable part. In this location, it is more sensitive to potential perturbations such as, for example, opportunist infections. In the third and last condition, the subject was already in the unstable phase due to a health issue, i.e. IBS. This can be observed in one of the patients from Durban \emph{et al.}'s study \cite{IBS}. In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost. It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see \ref{supfig:HLS_xWSummary}).

>>>>

By solving Langevin's differential equation, we can obtain a phase diagram where each microbiota sample can be placed according to its fitness and variability into one of two phases, according to the ranking stability of the system. As we can see in the phase--space in Figure \ref{fig:main3}, three different conditions that could occur are shown. First, we could have a healthy microbiota with some fluctuations, as shown by one of the subjects of Caporaso \emph{et al.}'s study \cite{moving}. Because this case would have good fitness, its temporal variability would not place the microbiota in the unstable phase of the diagram. Second, we have a subject from the study by Dethlefsen and Relman \cite{antibiotic} who was perturbed twice by an antibiotic intake. His microbiota underwent sufficient change so as to lose its stability, and hence be placed in the unstable part. In this location, it is more sensitive to potential perturbations such as, for example, opportunist infections. In the third and last condition, the subject was already in the unstable phase due to a health issue, i.e. IBS. This can be observed in one of the patients from Durban \emph{et al.}'s study \cite{IBS}. In addition, it was shown that this subject's health status improved during the time the experiment was carried out, implying that his microbiota also recovered the stability it had lost. It is interesting to note that in the subject's health from the study made by David \emph{et al.} \cite{hostlife} who suffered a Salmonella infection during the experiment, there was a significant shift in variability and a final recovery from the perturbed state (see Supplementary Figure  \ref{supfig:HLS_xWSummary}).
